---
title: "R Notebook"
output: html_notebook
---



# DDG2P: The relation between the number of phenotypes and DDD categories of mutation consequences


## 1. Load DDG2P dataset


```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggrepel)

DDG2P_hpo <- read.table("table.hpo_obo_20190924.txt", sep='\t', header = T)
DDG2P_pLI <- read.delim("DDG2P_24_9_2019.with_pLI.txt")
```


## 2. Visualization


I separated the phenotypes in each disease and changed its form into an easier form for calculation.
```{r}
ddg2p <- DDG2P_pLI %>%
  separate_rows(phenotypes, sep = ';') %>%
  na.omit()
ddg2p

nphenotypes <- as.numeric(as.factor(gsub("HP:0", "", ddg2p$phenotypes)))
nphenotypes

ddg2p <- mutate(ddg2p, phenotypes = nphenotypes)
```


Then, I selected 3 columns out of the whole data to make it easier for me to analyze.
```{r}
ddg2p %>%
  select('DDD.category', 'mutation.consequence', 'phenotypes')
```


I used the function `group_by` to group the table into `DDD.category` and `mutation.consequence`. Then I counted the number of phenotypes for the grouped form.
```{r}
ddg2p %>%
  select('DDD.category', 'mutation.consequence', 'phenotypes') %>%
  group_by(DDD.category, mutation.consequence) %>%
  count(name = 'no.phenotype')
```


I assigned the table to `b`.
```{r}
b <- ddg2p %>%
  select('DDD.category', 'mutation.consequence', 'phenotypes') %>%
  group_by(DDD.category, mutation.consequence) %>%
  count(name = 'no.phenotype')
b
```


```{r}
b %>% ggplot(aes(DDD.category, no.phenotype)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(mutation.consequence~.) +
  labs(title = 'Number of Phenotypes by DDD category and Mutation Consequence' , x = 'DDD category', y = 'Number of Phenotypes')
```


## 3. Modification


plot 1
```{r}
b %>% ggplot(aes(DDD.category, no.phenotype, fill = DDD.category, label = no.phenotype)) +
  geom_bar(stat = "identity") +
  geom_label_repel(size = 1.5) +
  scale_fill_discrete(name = 'DDD category') +
  coord_flip() +
  facet_wrap(mutation.consequence~.) +
  labs(title = 'Number of Phenotypes by DDD category and Mutation Consequence' , x = 'DDD category', y = 'Number of Phenotypes')
```

- I have added the labels to each values to make sure the small values could still be shown it was there.
- It was still difficult to precisely identify all values just by using `geom_text`. So I have replaced it with `geom_label_repel` in order for it to be distinguished easily.
- To make an improvement on the opinion that the plot looks too simple in black and white, I  have used the `fill = ` function to present different colors for each DDD category.


plot 2
```{r}
b %>% ggplot(aes(DDD.category, no.phenotype, fill = mutation.consequence, label = no.phenotype)) +
  geom_bar(stat = "identity") +
  theme(legend.key.width = unit(0.5, 'cm'), legend.position = 'bottom', legend.text = element_text(size=5), legend.title = element_text(size=7))+
  geom_label_repel(size = 2) +
  scale_fill_discrete(name = 'Mutation Consequence') +
  coord_flip() +
  labs(title = 'Number of Phenotypes by DDD category and Mutation Consequence' , x = 'DDD category', y = 'Number of Phenotypes')
```

- One opinion said that it wasn't nessesary to use `facet_wrap` for the plot. So I have changed the plot form into a barplot that is stacked with mutation consequences and sorted into DDD categories.
- The values for this plot were also very close to each other so I have used `geom_label_repel` to label numbers for each values.
- The legend was initially on the right side of the plot but I moved it to the bottom of the plot so that there were more space for the plot.



